---
  title: "No Perfect Collinearity"
output: 'pdf_document'  
classoption: landscape
fontsize: 12pt
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(wooldridge)
library(car)
library(lmtest)
library(sandwich)
library(stargazer)
```

## Check if the input variables inform each other

The regression we are analysing is the following: $ln(views)=β_0 + β_1 rate + β_3 length$

Where the input, $rate$, is a representation of the average rating of the video and $length$ is the video duration in seconds.

The risk of collinearity here is $length$ and $rate$ could be related to one another.

```{r Create the model, message=FALSE, echo=FALSE}
youtube_data <- read.table(file = "videos.txt", sep = "\t", header = TRUE)
youtube_model = lm(views ~ rate + length, data = youtube_data)
```
As we can see from the below results which reflect the coefficients of the model inputs, there is no sign of perfect collinearity since both variables have not been dropped. 

```{r Reveal the coefficients}
youtube_model$coefficients
```