---
title: "Chicago Public Schools Mathematics and Literacy Skills"
author: 'Austen Lowitz, Annie DeForge, William Teng'
output:
  pdf_document:
    toc: yes
  output:
  bookdown::pdf_document2:
    toc: yes
    number_sections: yes
  word_document:
    toc: yes
  fig_caption: yes
  includes:
    in_header: my_header.tex
geometry: "left=0.35in,right=0.35in,top=1in,bottom=1in"
header-includes:
 \usepackage{wrapfig}
 \usepackage{float}
 \floatplacement{figure}{H}
---

```{=tex}
\newpage
\setcounter{page}{1}
```
```{r load packages and set options, include=FALSE}
library(tidyverse)
library(magrittr)
library(stargazer)
library(sandwich)
library(lmtest)
library(dplyr)
library(ggplot2)
library(readr)
library(GGally)
library(reshape2)
library(car)
library(gridExtra)

theme_set(theme_bw())

options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(echo=FALSE, message=FALSE)
```

```{r load data, echo=FALSE, message = FALSE, include=FALSE}
#setwd("C:/Users/Annie's Computer/Documents/DS 203/lab 2/") 

# Set columns to focus on
school_columns <- c("Student.Performance.Reading.and.Math",
                    "Growth.Overall.Reading.and.Math",
                    "Student.Attendance.2012...Percent",
                    "Misconducts.Resulting.in.Suspensions.2012...Percent",
                    "Average.Days.of.Suspension.2012",
                    "Teacher.Attendance.2012...Percent",
                    "Involved.Families",
                    "Supportive.Environment",
                    "Safety",
                    "Effective.Leaders",
                    "Ambitious.Instruction",
                    "Collaborative.Teachers",
                    "Longitude",
                    "Latitude",
                    "Years.on.Probation",
                    "School.Track")


# Read in initial data : 460 Records
my_data <- read.csv("chicago-public-schools-elementary-school-progress-report-card-2012-2013-1.csv")[, school_columns] 


# Filter out NULLs : 416 Records
my_data <- my_data[complete.cases(my_data[, school_columns]), ]

# Filter out "Not Enough Data" : 282 Records
my_data <- subset(my_data, Involved.Families != "Not Enough Data"
                         & Supportive.Environment != "Not Enough Data"
                         & Safety != "Not Enough Data"
                         & Effective.Leaders != "Not Enough Data"
                         & Ambitious.Instruction != "Not Enough Data"
                         & Collaborative.Teachers != "Not Enough Data")

# Convert Likert variables into ascending scales (1 = Lowest, N = Highest)
my_data$Involved.Families <- factor(my_data$Involved.Families, levels = c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"), labels = c(1, 2, 3, 4, 5))

my_data$Supportive.Environment <- factor(my_data$Supportive.Environment, levels = c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"), labels = c(1, 2, 3, 4, 5))

my_data$Safety <- factor(my_data$Safety, levels = c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"), labels = c(1, 2, 3, 4, 5))

my_data$Effective.Leaders <- factor(my_data$Effective.Leaders, levels = c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"), labels = c(1, 2, 3, 4, 5))

my_data$Ambitious.Instruction <- factor(my_data$Ambitious.Instruction, levels = c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"), labels = c(1, 2, 3, 4, 5))

my_data$Collaborative.Teachers <- factor(my_data$Collaborative.Teachers, levels = c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"), labels = c(1, 2, 3, 4, 5))

# Convert numeric columns into numeric
my_data$Average.Days.of.Suspension.2012 <- as.numeric(my_data$Average.Days.of.Suspension.2012)  

# my_data$Involved.Families <- as.numeric(my_data$Involved.Families)
# 
# my_data$Supportive.Environment <- as.numeric(my_data$Supportive.Environment)
# 
# my_data$Safety <- as.numeric(my_data$Safety)
# 
# my_data$Effective.Leaders <- as.numeric(my_data$Effective.Leaders)
# 
# my_data$Ambitious.Instruction <- as.numeric(my_data$Ambitious.Instruction)
# 
# my_data$Collaborative.Teachers <- as.numeric(my_data$Collaborative.Teachers)

my_data$Growth.Overall.Reading.and.Math <- as.numeric(gsub("%", "", my_data$Growth.Overall.Reading.and.Math)) / 100

my_data$Student.Attendance.2012...Percent <- as.numeric(gsub("%", "", my_data$Student.Attendance.2012...Percent)) / 100

my_data$Teacher.Attendance.2012...Percent <- as.numeric(gsub("%", "", my_data$Teacher.Attendance.2012...Percent)) / 100
                    
my_data$Misconducts.Resulting.in.Suspensions.2012...Percent <- as.numeric(gsub("%", "", my_data$Misconducts.Resulting.in.Suspensions.2012...Percent)) / 100

my_data$Student.Performance.Reading.and.Math <- as.numeric(gsub("%", "", my_data$Student.Performance.Reading.and.Math)) / 100

# Drop nulls in dependent variable. For some reason the code above did not catch all nulls (don't know why)
my_data <- my_data[complete.cases(my_data$Growth.Overall.Reading.and.Math), ]
my_data <- my_data[complete.cases(my_data$Student.Performance.Reading.and.Math), ]

# Changing the name of columns because they are too long for the stargazer
my_data <- my_data %>%
  rename(Suspensions.Percent = Misconducts.Resulting.in.Suspensions.2012...Percent)
my_data <- my_data %>%
  rename(Growth.Reading.and.Math = Growth.Overall.Reading.and.Math)
my_data <- my_data %>%
  rename(Student.Attendance = Student.Attendance.2012...Percent)
my_data <- my_data %>%
  rename(Suspension.Days = Average.Days.of.Suspension.2012)
my_data <- my_data %>%
  rename(Teacher.Attendance = Teacher.Attendance.2012...Percent)

# Create new dependent variable: 
my_data$Student.Success <- (my_data$Student.Performance.Reading.and.Math + my_data$Growth.Reading.and.Math) / 2

```

# Introduction

In today's educational landscape, data-driven decision-making has emerged as an integral tool for enhancing the student experience. Leveraging the power of data allows educational stakeholders to identify critical factors that contribute to a student's academic success, allowing for effective interventions and informed policy adjustments[^1].

[^1]: Mandinach, E. B., & Gummer, E. (n.d.). Data-driven decision making: Components of the ... - sage journals. <https://journals.sagepub.com/doi/abs/10.1177/016146811511700402>

The Chicago Public School District is amongst the largest and most diverse urban school districts in the US. With that comes unique challenges and opportunities in applying data to drive academic improvement. Currently, only 21% of students had proficient ratings on the statewide reading exam, and only 17% had proficient ratings in math[^2]. Recognizing this potential, our analysis seeks to uncover evidence-based actionable insights to enhance student performance for Chicago's elementary school students.

[^2]: Senior. (2023, June 29). Explore chicago public schools. Niche. <https://www.niche.com/k12/d/chicago-public-schools-il/>

Through this analysis, we aim to provide Chicago elementary school superintendents with a robust framework for developing multi-year roadmaps to enhance academic success. This can include changes from staffing and leadership decisions, school security code changes, and teaching staff rostering to disciplinary policies. The end result will be a more responsive, equitable, and effective educational system that empowers students to thrive academically. This work contributes to the broader movement towards data-informed education, embodying the shared vision of educators, policymakers, and researchers alike to harness the transformative potential of data for the betterment of our schools and communities.

This study estimates the average proportion of elementary school students who are meeting progress and proficiency targets in Chicago. The data shows details on the status of the school and the quality of the classroom environment. We propose a set of regression models to show the learning environment variables that are significant to predicting student success. 

# Data and Methodology

The data for this study is from the 2012 School Progress Report Card for the elementary schools in the Chicago Public School System. It was made public by the city of Chicago. Each row represents an elementary school in the school district. The dataset had 460 rows, however, there were schools that had missing values or had a "not enough data record" in our variables of interest. There seemed to be no obvious pattern in the schools that had missing values, so we chose to exclude these from our analysis, which left us with 218 observations.

Average daily student attendance was one of the main variables of interest provided in the dataset. Studies have shown that student attendance is one of the most important factors in student success in the classroom[^3]. Motivation in general has been shown to be very important for learning, and attendance is a key measure of a student's academic motivation[^4]. Additionally, this dataset also contains a grade for several learning environment variables. Schools were graded on a 5 point scale from "very weak" to "very strong" on how well the school involved families, how supportive the environment was, how safe the school was, if the leadership was focused, if the instruction was focused and challenging, and how well the teachers worked together. Additionally, there was an average daily attendance variable for teachers. These student motivation and learning environment variables were the ones that we were primarily interested in to understand which ones were significantly associated with academic success. We are interested in using a regression model to evaluate which of these classroom experience variables have significant coefficients to explain the variability in academic success.

[^3]: Every School Day Counts: The Forum Guide to collecting and using attendance data. National Center for Education Statistics (NCES) Home Page, a part of the U.S. Department of Education. (n.d.). <https://nces.ed.gov/pubs2009/attendancedata/chapter1a.asp>.

[^4]: Davis, L. (2015). STUDENT ATTENDANCE: A SCHOOL'S INTERVENTION TO INCREASE ATTENDANCE. [online] Available at: <https://www.nwmissouri.edu/library/researchpapers/2015/Davis,%20Luke.pdf> [Accessed 7 Aug. 2023].

We also wanted to control for other factors that impact the school that were not directly related to what was in the classroom. To do so, we incorporated an indicator variable showing if the school was on a regular schedule or a shortened summer schedule to reduce learning loss over summer break, the number of years the school had been on probation, disciplinary policy (the number of misconducts resulting in suspension and average days of suspension), and location (longitude and latitude) as variables to consider in our model.

We operationalized student success to incorporate two metrics: the proportion of students at the school who met or exceeded the national average on reading and math test, and the proportion of students who improved the expected amount between the fall and spring tests. We averaged these two proportions to create a combined metric on student success, which we used as our outcome variable in our models.

```{r Distribution and Descriptive Stats, echo=FALSE, include=FALSE}


# histogram of attendence variable
hist(my_data$Student.Attendance)
hist(my_data$Teacher.Attendance)
# histogram of discipline variables
hist(my_data$Suspensions.Percent)
hist(my_data$Suspension.Days)

# histogram of likert variables
# hist(my_data$Effective.Leaders)
# hist(my_data$Involved.Families)
# hist(my_data$Safety)
# hist(my_data$Supportive.Environment)
# hist(my_data$Ambitious.Instruction)
# hist(my_data$Collaborative.Teachers)

# plots for attendance vs growth
plot(my_data$Student.Attendance, my_data$Student.Success)
plot(my_data$Teacher.Attendance, my_data$Student.Success)

# plots for discipline vs growth
plot(my_data$Suspension.Days, my_data$Student.Success)
plot(my_data$Suspensions.Percent, my_data$Student.Success)

# tables of average growth across likert categories
tapply(my_data$Student.Success, my_data$Safety, mean)
tapply(my_data$Student.Success, my_data$Involved.Families, mean)
tapply(my_data$Student.Success, my_data$Supportive.Environment, mean)
tapply(my_data$Student.Success, my_data$Effective.Leaders, mean)
tapply(my_data$Student.Success, my_data$Ambitious.Instruction, mean)
tapply(my_data$Student.Success, my_data$Collaborative.Teachers, mean)
```

```{r Location eda, echo=FALSE, include=FALSE}
#location exploration
# my_data$Student.Success <- as.numeric(gsub("%", "", my_data$Student.Success)) * 100

plot(my_data$Latitude, my_data$Longitude)
plot(my_data$Latitude, my_data$Student.Success)
plot(my_data$Longitude, my_data$Student.Success)
plot(my_data$Longitude*my_data$Latitude, my_data$Student.Success)

p<- ggplot(my_data, aes(x = Latitude,
               y = Longitude,
               color = Student.Success)) + geom_point()
```

```{r learning environment plot, echo=FALSE, fig.align='center'}

p1 <- ggplot(my_data, aes(x = Student.Attendance, y = Student.Success)) +
  geom_point(color = "blue", alpha = 0.5) +  # Scatter plot with blue points
  geom_smooth(method = "lm", color = "red", se = FALSE) + # Linear regression line in red
  labs(
    title = "Student Attendance vs. Student Success",
    x = "Student Attendance",
    y = "Student Success"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10)
  )

grade <- rep(c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"), 6)
grade <- factor(grade, levels = c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"))

variable <- c(rep("Involved Families", 5), rep("Supportive Environment", 5), rep("Safety", 5), rep("Effective Leaders", 5), rep("Ambitious Instructions", 5), rep("Collaborative Teachers", 5))

mean <- tapply(my_data$Student.Success, my_data$Involved.Families, mean)
mean <- c(mean, tapply(my_data$Student.Success, my_data$Supportive.Environment, mean))
mean <- c(mean, tapply(my_data$Student.Success, my_data$Safety, mean))
mean <- c(mean, tapply(my_data$Student.Success, my_data$Effective.Leaders, mean))
mean <- c(mean, tapply(my_data$Student.Success, my_data$Ambitious.Instruction, mean))
mean <- c(mean, tapply(my_data$Student.Success, my_data$Collaborative.Teachers, mean))

d <- data.frame(variable, grade, mean)

p2 <- ggplot(d, aes(fill = variable, y = mean, x = grade)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(
    title = "Mean Student Success Pct by Learning Environment Variable",
    y = "Mean Student Success Pct" # Naming the y-axis
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10)
  )
grid.arrange(p1, p2)
```

Our exploratory plots showed a roughly linear relationship between our variables of interest and student success. The structure of our model was the following:

$$\hat{\text{Student Success}} = \beta_0 + \beta_1 \cdot \text{Student Attendance} + \text{Y}\gamma + \text{Z}\delta$$ Y is a row vector of learning environment variables and $\gamma$ is a column vector of the coefficients. Z is a row vector of the additional control covariates and $\delta$ is a column vector of the coefficients.

Because of the importance of student attendance, we wanted to run a naive model with just this variable. We ran a wald test for nested models, testing one variable at a time starting with the learning environment variables as they appeared in the dataset to identify the significant learning environment variables and did the same with the other covariates.

```{r Correlation Matrix, echo=FALSE, include=FALSE}
# Exclude the specific columns from the subset
subset <- my_data[, setdiff(names(my_data), c("Student.Success", "Student.Performance.Reading.and.Math", "Growth.Reading.and.Math"))]

# Only include numeric columns
subset <- subset[, unlist(lapply(subset, is.numeric))]

# Compute the correlation between each predictor variable and the dependent variable
correlations <- cor(subset, my_data$Student.Success)

# Load the required packages
library(ggplot2)
library(reshape2)

# Convert the corr matrix to a df for viz
cor_df <- melt(correlations)

# Reorder the levels of the Var1 variable based on the correlation with the dependent variable
cor_df$Var1 <- reorder(cor_df$Var1, cor_df$value, FUN = function(x) -abs(x))

# Create a corr heatmap
heatmap_plot <- ggplot(cor_df, aes(x = Var1, y = Var2, fill = value, label = round(value, 2))) +
  geom_tile() +
  geom_text(color = "black", size = 6, vjust = 1) +  # Add correlation value as text
  scale_fill_gradient(low = "yellow", high = "purple") +
  labs(title = "Correlation of Independent Vars w/ Respect to Dependent Var") +
  labs(x = "", y = "") +  # Remove axis labels for Var1 and Var2
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  coord_fixed()

heatmap_plot <- heatmap_plot + labs(y = "Student.Success")

heatmap_plot
```

```{r, echo=FALSE, include=FALSE}
heatmap_plot
```

```{r, echo=FALSE, include=FALSE}
dim(my_data)
```

```{r Create some models, echo=FALSE}
# Naive Model: A few input variables
model_naive <- lm(Student.Success ~ Student.Attendance, data = my_data)

# Complex Model: All the variables
model_complex <- lm(Student.Success ~
                    Student.Attendance +
                    Suspensions.Percent +
                    Suspension.Days +
                    Teacher.Attendance +
                    Involved.Families +
                    Supportive.Environment +
                    Safety +
                    Effective.Leaders +
                    Ambitious.Instruction +
                    Collaborative.Teachers +
                    Longitude +
                    Latitude + 
                    School.Track +
                    Years.on.Probation, data = my_data)

learning_environment <- lm(Student.Success ~ Student.Attendance +
                        Teacher.Attendance +
                        Involved.Families +
                        Supportive.Environment +
                        Safety, data = my_data)

# Significant predictors plus other significant variables from complex model unrelated to learning_environment
controls <- lm(Student.Success ~ Student.Attendance +
                            Teacher.Attendance +
                            Involved.Families +
                            Supportive.Environment +
                            Safety +
                            School.Track +
                            Years.on.Probation, data = my_data)
```

```{r naive summary, echo=FALSE, include=FALSE}
summary(model_naive)
```

```{r complex summary, echo=FALSE, include=FALSE}
summary(model_complex)
```

```{r nested model test with wald test, echo=FALSE, include = FALSE}
# Teacher Attendance
mod2 <- lm(Student.Success~Student.Attendance + Teacher.Attendance, data = my_data)

waldtest(model_naive, mod2, vcov=vcovHC(mod2, type="HC0")) # significant

# Involved Families
mod3 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families , data = my_data)

waldtest(mod2, mod3, vcov=vcovHC(mod3, type="HC0")) # significant

# Supportive Environment
mod4 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment, data = my_data)

waldtest(mod3, mod4, vcov=vcovHC(mod4, type="HC0")) # significant

# Safety
mod5 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety, data = my_data)

waldtest(mod4, mod5, vcov=vcovHC(mod5, type="HC0")) # significant

# Effective Leaders
mod6 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Effective.Leaders, data = my_data)

waldtest(mod5, mod6, vcov=vcovHC(mod6, type="HC0")) # not significant

# Ambitious Instruction
mod7 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Ambitious.Instruction, data = my_data)

waldtest(mod5, mod7, vcov=vcovHC(mod7, type="HC0")) # not significant

# Collaborative Teachers
mod8 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Collaborative.Teachers, data = my_data)

waldtest(mod5, mod8, vcov=vcovHC(mod8, type="HC0")) # not significant

# Discipline Variables
mod9 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Suspension.Days + Suspensions.Percent, data = my_data)

waldtest(mod5, mod9,  vcov=vcovHC(mod9, type="HC0")) # not significant

# years on probation
mod10 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Years.on.Probation, data=my_data)

waldtest(mod5, mod10,  vcov=vcovHC(mod10, type="HC0")) # significant

# latitude
mod11 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Years.on.Probation + Latitude, data=my_data)

waldtest(mod10, mod11,  vcov=vcovHC(mod11, type="HC0")) # not significant

# longitude
mod12 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Years.on.Probation + Longitude, data=my_data)

waldtest(mod10, mod12,  vcov=vcovHC(mod12, type="HC0")) # not significant

# School.Track
mod13 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Years.on.Probation + School.Track, data=my_data)

waldtest(mod10, mod13,  vcov=vcovHC(mod13, type="HC0")) # not significant

```

# Results

Table 1 below compares four models: our "Naive" model with just one predictor variable, $\text{Student Attendance}$, a model containing our learning environment variables, a model with significant learning environment variables plus significant controls, like $\text{School Track}$ and $\text{Years on Probation}$. Last but not least, we have a "Complex" model which includes all our predictor variables.

Across all models, the beta coefficient on $\text{Student Attendance}$ was highly significant (p<0.01), indicating a strong positive relationship with student success. The coefficient point estimate decreases as more features are added to the model, ranging from 2.92 in the Naive model to 1.08 in the Complex model. This shows how the beta coefficient is pushed down by the inclusion of other environment and control variables.

```{r Stargazer Table, message=FALSE, echo=FALSE, results='asis'}

stargazer(model_naive, learning_environment, controls, model_complex,
          title = "Model Comparison",
          header = FALSE,
          table.placement = "!ht",
          column.labels = c("Naive", "Learning Environment", "Controls", "Complex"),
          dep.var.caption = "Regression Table",
          dep.var.labels.include = FALSE,
          omit = c("Effective.Leaders", "Ambitious.Instruction", "Collaborative.Teachers", "Involved.Families", "Supportive.Environment", "Safety"),
                    covariate.labels = c("Student Attendance", " Suspension Percent", "Suspension Days", "Teacher Attendance", "Longitude", "Latitude", "School Schedule (Track E)", "Years on Probation", "Constant"),

          add.lines = list(
            c("Involved Families", "", "\\checkmark", "\\checkmark","\\checkmark"),
            c("Supportive Environment", "", "\\checkmark", "\\checkmark", "\\checkmark"),
            c("Safety", "", "\\checkmark","\\checkmark","\\checkmark"),
            c("Effective.Leaders", "","","","\\checkmark"),
            c("Ambitious.Instruction", "","","","\\checkmark"),
            c("Collaborative.Teachers", "","","","\\checkmark"),
            "\\hline"
          ), 
          single.row = TRUE,  # Combine into a single row
          type = "latex", # Output in LaTeX format
          digits=2,
          font.size="small",
          notes.append = FALSE,
          notes = "\\parbox[t]{7cm}{School Schedule (Track E) is an indicator variable for having a shortened summer schedule. The categorical covariates each have 5 levels, a grade on the category from 'very weak' to 'very strong'.}"
          ) 
```

To illustrate the scale of these effects, consider a hypothetical school with a Student Attendance rate of 90%. Applying the learning environment model with controls, a school that experienced a increase in student attendance by 10 percentage points would be predicted to experience a 12 percentage point increase in the average proportion of students who are meeting or exceeding the national average or meeting expected growth rate in a school year.
Although $\text{Teacher Attendance}$ was significant in the nested model test when we added it as a second variable, once other significant variables were added in, it lost it's significance. This was due to the standard error being fairly high. 

Two variables, $\text{School Schedule (Track E)}$ and $\text{Years on Probation}$, appear to have a significant negative relationship on student success in the Controls and Complex models. $\text{Years on Probation}$ had a small effect size, but with a very small standard error and it was very robust to different model specifications. A school being on a reduced summer schedule (Track E) was also negatively associated with $\text{Student Success}$. There may be an element of reverse causality here in that a school where students are struggling more is more likely to try a different summer schedule to reduce school break learning loss.   

Several variables, including $\text{Suspensions Percent}$, $\text{Suspension Days}$, $\text{Longitude}$, and $\text{Latitude}$, were included in the Complex model but did not show statistical significance. This is due to some extent on our operationalization of $\text{Student Success}$. For instance, there is a significant North-South divide on socio-economic status in Chicago, so longitude had a strong linear relationship with national performance on math and reading tests, but an insignificant relationship with growth on math and reading tests. 

From examining the Complex model, we can see some of the negative effects of including insignificant variables. In the complex model, the confidence on $\text{Teacher Attendance}$ is likely very overestimated since it increases dramatically from the learning environment with controls model. Additionally, in the complex model, the coefficients of the dummy variables from improving the grade $\text{Supportive Environment}$ flips from a positive association with $\text{Student Success}$ to a negative one. 

Considering the different model specifications, the Learning Environment with controls is the best model because it can give estimates for the classroom related variables that teachers and superintendents will be interested to learn about, while controlling for other school factors, without creating the interpretability problems that would arise from using the less parsimonious complex model. 

# Limitations

Given our large sample size (218 observations), we can utilize a multiple regression model to assess the key drivers on student success, assuming that our data meets the following statistical limitations: The data must be Independent and Identically Distributed (IID), and there must exist no perfect collinearity between features.

Each row in our dataframe represents a unique elementary school in Chicago. Because the data includes the entire population of elementary schools in Chicago rather than a specific sample, we have reduced concerns related to sampling bias since every school is represented in the dataset. Since we are analyzing the entire population of elementary schools in Chicago, we assume that the schools operate under similar conditions and follow similar educational standards and policies. This supports the notion that our observations are drawn from the same distribution, therefore meeting the IID assumption. That said, our scope is limited towards only being able to make generalizations about Chicago public schools.

We examined the relationships between independent variables through a correlation matrix and found no correlations near 1, indicating no perfect collinearity. We did not observe any correlations that were high enough to be a concern. Additionally, we examined the VIF of all of the possible variables together (the complex model) and saw that all of the variables were below 5, which shows that multicollinearity is not a concern for this data. 

Since our data is IID and has no perfect collinearity, we pass all of our large linear model assumptions and therefore rule out any concerns for using a multiple regression model for this analysis.

Regarding structural limitations, one key omitted variable is school funding, which could have significant implications for our model's accuracy. School funding is expected to be positively correlated with our target variable, $\text{Student Success}$. Moreover, more school funding would likely enhance some of our learning environment variables like safety and student attendance. By having school funding as an omitted variable, we may be inadvertently attributing more significance to our existing predictor variables, thus overestimating their impact. Including data on school funding would therefore drive the main effect towards zero, reducing the likelihood of a type I error.

# Conclusion

Through our analysis, we can conclude that for elementary students in Chicago public schools, student's success is significantly associated with their school attendance, school status, and learning environment, including variables like $\text{Involved Families}$, $\text{Supportive Environment}$, and $\text{Safety}$. 

Since student success can benefit society in many ways, from increasing tax revenue to creating a more vibrant society[^5], it is crucial for superintendents to use this information to align priorities when developing future road maps to drive student success. As a starting point, superintendents can focus on increasing student attendance, which had the largest coefficient amongst our significant predictor variables.  Crucially, by identifying significant relationships, further randomized experimental research can be done on the association that were found to be signficant through out our investigation in order to establish what the causal relationship is between the learning environment and student's academic success.

[^5]: How do college graduates benefit society at large?. APLU. (2023, March 1). <https://www.aplu.org/our-work/4-policy-and-advocacy/publicuvalues/societal-benefits/>

That said, addressing the challenge of student attendance requires a multifaceted approach; one that goes beyond simply boosting attendance numbers. It calls for initiatives that foster a safer school environment, cultivate family engagement, and address underlying issues such as safety concerns that may be hindering attendance in the first place. These efforts can resonate beyond attendance, potentially reducing the number of years a school is on probation---another significant predictor of student success in our study.

In conclusion, our study serves as a testament to the power of data-driven decision-making in education. By delving into the complex interplay of factors that contribute to student success, we have provided actionable insights that can be harnessed to foster a more enriching educational experience for elementary school students in Chicago.
