---
title: "Chicago Public Schools Mathematics and Literacy Skills"
author: 'Austen Lowitz, Annie DeForge, William Teng'
output:
  pdf_document:
    toc: yes
  output:
  bookdown::pdf_document2:
    toc: yes
    number_sections: yes
  word_document:
    toc: yes
  fig_caption: yes
  includes:
    in_header: my_header.tex
geometry: "left=0.5in,right=0.5in,top=0.5in,bottom=0.5in"
header-includes:
 \usepackage{wrapfig}
 \usepackage{float}
 \floatplacement{figure}{H}
---

```{r}
tinytex::install_tinytex()
```


```{=tex}
\newpage
\setcounter{page}{1}
```
```{r load packages and set options, include=FALSE}
library(tidyverse)
library(magrittr)
library(stargazer)
library(sandwich)
library(lmtest)
library(dplyr)
library(ggplot2)
library(readr)
library(GGally)
library(reshape2)
library(car)
library(gridExtra)

theme_set(theme_bw())

options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(echo=FALSE, message=FALSE)
```

```{r load data, echo=FALSE, message = FALSE, include=FALSE}
#setwd("C:/Users/Annie's Computer/Documents/DS 203/lab 2/") 

# Set columns to focus on
school_columns <- c("Student.Performance.Reading.and.Math",
                    "Growth.Overall.Reading.and.Math",
                    "Student.Attendance.2012...Percent",
                    "Misconducts.Resulting.in.Suspensions.2012...Percent",
                    "Average.Days.of.Suspension.2012",
                    "Teacher.Attendance.2012...Percent",
                    "Involved.Families",
                    "Supportive.Environment",
                    "Safety",
                    "Effective.Leaders",
                    "Ambitious.Instruction",
                    "Collaborative.Teachers",
                    "Longitude",
                    "Latitude",
                    "Years.on.Probation",
                    "School.Track")


# Read in initial data : 460 Records
my_data <- read.csv("chicago-public-schools-elementary-school-progress-report-card-2012-2013-1.csv")[, school_columns] 


# Filter out NULLs : 416 Records
my_data <- my_data[complete.cases(my_data[, school_columns]), ]

# Filter out "Not Enough Data" : 282 Records
my_data <- subset(my_data, Involved.Families != "Not Enough Data"
                         & Supportive.Environment != "Not Enough Data"
                         & Safety != "Not Enough Data"
                         & Effective.Leaders != "Not Enough Data"
                         & Ambitious.Instruction != "Not Enough Data"
                         & Collaborative.Teachers != "Not Enough Data")

# Convert Likert variables into ascending scales (1 = Lowest, N = Highest)
my_data$Involved.Families <- factor(my_data$Involved.Families, levels = c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"), labels = c(1, 2, 3, 4, 5))

my_data$Supportive.Environment <- factor(my_data$Supportive.Environment, levels = c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"), labels = c(1, 2, 3, 4, 5))

my_data$Safety <- factor(my_data$Safety, levels = c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"), labels = c(1, 2, 3, 4, 5))

my_data$Effective.Leaders <- factor(my_data$Effective.Leaders, levels = c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"), labels = c(1, 2, 3, 4, 5))

my_data$Ambitious.Instruction <- factor(my_data$Ambitious.Instruction, levels = c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"), labels = c(1, 2, 3, 4, 5))

my_data$Collaborative.Teachers <- factor(my_data$Collaborative.Teachers, levels = c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"), labels = c(1, 2, 3, 4, 5))

# Convert numeric columns into numeric
my_data$Average.Days.of.Suspension.2012 <- as.numeric(my_data$Average.Days.of.Suspension.2012)  

# my_data$Involved.Families <- as.numeric(my_data$Involved.Families)
# 
# my_data$Supportive.Environment <- as.numeric(my_data$Supportive.Environment)
# 
# my_data$Safety <- as.numeric(my_data$Safety)
# 
# my_data$Effective.Leaders <- as.numeric(my_data$Effective.Leaders)
# 
# my_data$Ambitious.Instruction <- as.numeric(my_data$Ambitious.Instruction)
# 
# my_data$Collaborative.Teachers <- as.numeric(my_data$Collaborative.Teachers)

my_data$Growth.Overall.Reading.and.Math <- as.numeric(gsub("%", "", my_data$Growth.Overall.Reading.and.Math)) / 100

my_data$Student.Attendance.2012...Percent <- as.numeric(gsub("%", "", my_data$Student.Attendance.2012...Percent)) / 100

my_data$Teacher.Attendance.2012...Percent <- as.numeric(gsub("%", "", my_data$Teacher.Attendance.2012...Percent)) / 100
                    
my_data$Misconducts.Resulting.in.Suspensions.2012...Percent <- as.numeric(gsub("%", "", my_data$Misconducts.Resulting.in.Suspensions.2012...Percent)) / 100

my_data$Student.Performance.Reading.and.Math <- as.numeric(gsub("%", "", my_data$Student.Performance.Reading.and.Math)) / 100

# Drop nulls in dependent variable. For some reason the code above did not catch all nulls (don't know why)
my_data <- my_data[complete.cases(my_data$Growth.Overall.Reading.and.Math), ]
my_data <- my_data[complete.cases(my_data$Student.Performance.Reading.and.Math), ]

# Changing the name of columns because they are too long for the stargazer
my_data <- my_data %>%
  rename(Suspensions.Percent = Misconducts.Resulting.in.Suspensions.2012...Percent)
my_data <- my_data %>%
  rename(Growth.Reading.and.Math = Growth.Overall.Reading.and.Math)
my_data <- my_data %>%
  rename(Student.Attendance = Student.Attendance.2012...Percent)
my_data <- my_data %>%
  rename(Suspension.Days = Average.Days.of.Suspension.2012)
my_data <- my_data %>%
  rename(Teacher.Attendance = Teacher.Attendance.2012...Percent)

# Create new dependent variable: 
my_data$Student.Success <- (my_data$Student.Performance.Reading.and.Math + my_data$Growth.Reading.and.Math) / 2

```

# Introduction

In today's educational landscape, data-driven decision-making has emerged as an integral tool for enhancing the student experience. Leveraging the power of data allows educational stakeholders to identify critical factors that contribute to a student's academic success, allowing for effective interventions and informed policy adjustments[^1].

[^1]: Mandinach, E. B., & Gummer, E. (n.d.). Data-driven decision making: Components of the ... - sage journals. <https://journals.sagepub.com/doi/abs/10.1177/016146811511700402>

The Chicago Public School District is amongst the largest and most diverse urban school districts in the US. With that comes unique challenges and opportunities in applying data to drive academic improvement. Currently, only 21% of students had proficient ratings on the statewide reading exam, and only 17% had proficient ratings in math[^2]. Recognizing this potential, our analysis seeks to uncover evidence-based actionable insights to enhance student performance for Chicago's elementary school students.

[^2]: Senior. (2023, June 29). Explore chicago public schools. Niche. <https://www.niche.com/k12/d/chicago-public-schools-il/>

Through this analysis, we aim to provide Chicago elementary school superintendents with a robust framework for developing multi-year roadmaps to enhance academic success. This can include changes from staffing and leadership decisions, school security code changes, and teaching staff rostering to disciplinary policies. The end result will be a more responsive, equitable, and effective educational system that empowers students to thrive academically. This work contributes to the broader movement towards data-informed education, embodying the shared vision of educators, policymakers, and researchers alike to harness the transformative potential of data for the betterment of our schools and communities.

# Data and Methodology

The data for this study is from the 2012 School Progress Report Card for the elementary schools in the Chicago Public School System, it was made public by the city of Chicago. Each row represents an elementary school in the school district. The dataset had 460 rows, however, there were schools that had missing values or had a "not enough data record" in our variables of interest. There seemed to be no obvious pattern in the schools that had missing values, so we chose to exclude these from our analysis, which left us with 218 observations. 

Average daily student attendance was one of the main variables of interest provided in the dataset. Studies have shown that student attendance is one of the most important factors in student success in the classroom[^3]. Motivation in general has been shown to be very important for learning, and attendance is a key measure of a student's academic motivation[^4]. Additionally, this dataset also contains a grade for several learning environment variables. Schools were graded on a 5 point scale from "very weak" to "very strong" on how well the school involved families, how supportive the environment was, how safe the school was, if the leadership was focused, if the instruction was focused and challenging, and how well the teachers worked together. Additionally, there was average daily attendance variable for teachers. These student motivation and learning environment variables were the ones that we were primarily interested in to understand which ones were significantly associated with academic success. We are interested in using a regression model to evaluate which of these classroom experience variables have significant coefficients to explain the variability in academic success. 

We also wanted to control for other factors that impact the school that were not directly related to what was in the classroom. To do so, we incorporated an indicator variable showing if the school was on a regular schedule or a shortened summer schedule to reduce learning loss over summer break, the number of years the school had been on probation, disciplinary policy (the number of misconducts resulting in suspension and average days of suspension), and location (longitude and latitude) as variables to consider in our model.

We operationalized student success to incorporate two metrics: they proportion of students at the school who met or exceeded the national average on reading and math test and the proportion of students who improved the expected amount between the fall and spring tests. We averaged these two proportions to create a combined metric on student success, which we used as our outcome variable in our models.

```{r Distribution and Descriptive Stats, echo=FALSE, include=FALSE}


# histogram of attendence variable
hist(my_data$Student.Attendance)
hist(my_data$Teacher.Attendance)
# histogram of discipline variables
hist(my_data$Suspensions.Percent)
hist(my_data$Suspension.Days)

# histogram of likert variables
# hist(my_data$Effective.Leaders)
# hist(my_data$Involved.Families)
# hist(my_data$Safety)
# hist(my_data$Supportive.Environment)
# hist(my_data$Ambitious.Instruction)
# hist(my_data$Collaborative.Teachers)

# plots for attendance vs growth
plot(my_data$Student.Attendance, my_data$Student.Success)
plot(my_data$Teacher.Attendance, my_data$Student.Success)

# plots for discipline vs growth
plot(my_data$Suspension.Days, my_data$Student.Success)
plot(my_data$Suspensions.Percent, my_data$Student.Success)

# tables of average growth across likert categories
tapply(my_data$Student.Success, my_data$Safety, mean)
tapply(my_data$Student.Success, my_data$Involved.Families, mean)
tapply(my_data$Student.Success, my_data$Supportive.Environment, mean)
tapply(my_data$Student.Success, my_data$Effective.Leaders, mean)
tapply(my_data$Student.Success, my_data$Ambitious.Instruction, mean)
tapply(my_data$Student.Success, my_data$Collaborative.Teachers, mean)
```

```{r Location eda, echo=FALSE, fig.keep='last'}
#location exploration
# my_data$Student.Success <- as.numeric(gsub("%", "", my_data$Student.Success)) * 100

plot(my_data$Latitude, my_data$Longitude)
plot(my_data$Latitude, my_data$Student.Success)
plot(my_data$Longitude, my_data$Student.Success)
plot(my_data$Longitude*my_data$Latitude, my_data$Student.Success)

p<- ggplot(my_data, aes(x = Latitude,
               y = Longitude,
               color = Student.Success)) + geom_point()
```

```{r learning environment plot, echo=FALSE}

p1 <- ggplot(my_data, aes(x = Student.Attendance, y = Student.Success)) +
  geom_point(color = "blue", alpha = 0.5) +  # Scatter plot with blue points
  geom_smooth(method = "lm", color = "red", se = FALSE) + # Linear regression line in red
  labs(
    title = "Student Attendance vs. Student Success",
    x = "Student Attendance",
    y = "Student Success"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )

grade <- rep(c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"), 6)
grade <- factor(grade, levels = c("Very Weak", "Weak", "Neutral", "Strong", "Very Strong"))

variable <- c(rep("Involved Families", 5), rep("Supportive Environment", 5), rep("Safety", 5), rep("Effective Leaders", 5), rep("Ambitious Instructions", 5), rep("Collaborative Teachers", 5))

mean <- tapply(my_data$Student.Success, my_data$Involved.Families, mean)
mean <- c(mean, tapply(my_data$Student.Success, my_data$Supportive.Environment, mean))
mean <- c(mean, tapply(my_data$Student.Success, my_data$Safety, mean))
mean <- c(mean, tapply(my_data$Student.Success, my_data$Effective.Leaders, mean))
mean <- c(mean, tapply(my_data$Student.Success, my_data$Ambitious.Instruction, mean))
mean <- c(mean, tapply(my_data$Student.Success, my_data$Collaborative.Teachers, mean))

d <- data.frame(variable, grade, mean)

p2 <- ggplot(d, aes(fill = variable, y = mean, x = grade)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(
    title = "Mean Student Success Pct by Learning Environment Variable",
    y = "Mean Student Success Pct" # Naming the y-axis
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )
grid.arrange(p1, p2)
```
Our exploratory plots showed a roughly linear relationship between our variables of interest and student success. The structure of our model was

$$\hat{\text{Student Success}} = \beta_0 + \beta_1 \cdot \text{Student Attendance} + \text{Y}\gamma + \text{Z}\delta$$
Y is a row vector of learning environment variables and $\gamma$ is a column vector of the coefficients. Z is a row vector of the additional control covariates and $\delta$ is a column vector of the coefficients.

Because of the importance of student attendance, we wanted to run a naive model with just this variable. We ran a wald test for nested models, testing one variable at a time starting with the learning environment variables as they appeared in the dataset to identify the significant learning environment variables and did the same with the other covariates. 

[^3]: Every School Day Counts: The Forum Guide to collecting and using attendance data. National Center for Education Statistics (NCES) Home Page, a part of the U.S. Department of Education. (n.d.). https://nces.ed.gov/pubs2009/attendancedata/chapter1a.asp.
[^4]: Davis, L. (2015). STUDENT ATTENDANCE: A SCHOOL’S INTERVENTION TO INCREASE ATTENDANCE. [online] Available at: https://www.nwmissouri.edu/library/researchpapers/2015/Davis,%20Luke.pdf [Accessed 7 Aug. 2023].


```{r Correlation Matrix, echo=FALSE, include=FALSE}
# Exclude the specific columns from the subset
subset <- my_data[, setdiff(names(my_data), c("Student.Success", "Student.Performance.Reading.and.Math", "Growth.Reading.and.Math"))]

# Only include numeric columns
subset <- subset[, unlist(lapply(subset, is.numeric))]

# Compute the correlation between each predictor variable and the dependent variable
correlations <- cor(subset, my_data$Student.Success)

# Load the required packages
library(ggplot2)
library(reshape2)

# Convert the corr matrix to a df for viz
cor_df <- melt(correlations)

# Reorder the levels of the Var1 variable based on the correlation with the dependent variable
cor_df$Var1 <- reorder(cor_df$Var1, cor_df$value, FUN = function(x) -abs(x))

# Create a corr heatmap
heatmap_plot <- ggplot(cor_df, aes(x = Var1, y = Var2, fill = value, label = round(value, 2))) +
  geom_tile() +
  geom_text(color = "black", size = 6, vjust = 1) +  # Add correlation value as text
  scale_fill_gradient(low = "yellow", high = "purple") +
  labs(title = "Correlation of Independent Vars w/ Respect to Dependent Var") +
  labs(x = "", y = "") +  # Remove axis labels for Var1 and Var2
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  coord_fixed()

heatmap_plot <- heatmap_plot + labs(y = "Student.Success")

heatmap_plot
```

```{r, echo=FALSE, include=FALSE}
dim(my_data)
```

```{r Create some models, echo=FALSE}
# Naive Model: A few input variables
model_naive <- lm(Student.Success ~ Student.Attendance, data = my_data)

# Complex Model: All the variables
model_complex <- lm(Student.Success ~
                    Student.Attendance +
                    Suspensions.Percent +
                    Suspension.Days +
                    Teacher.Attendance +
                    Involved.Families +
                    Supportive.Environment +
                    Safety +
                    Effective.Leaders +
                    Ambitious.Instruction +
                    Collaborative.Teachers +
                    Longitude +
                    Latitude + 
                    School.Track +
                    Years.on.Probation, data = my_data)

learning_environment <- lm(Student.Success ~ Student.Attendance +
                        Teacher.Attendance +
                        Involved.Families +
                        Supportive.Environment +
                        Safety +
                        Effective.Leaders +
                        Ambitious.Instruction +
                        Collaborative.Teachers, data = my_data)

# Significant predictors plus other significant variables from complex model unrelated to learning_environment
controls <- lm(Student.Success ~ Student.Attendance +
                            Teacher.Attendance +
                            Involved.Families +
                            Supportive.Environment +
                            Safety +
                            School.Track +
                            Years.on.Probation, data = my_data)
```

```{r naive summary, echo=FALSE, include=FALSE}
summary(model_naive)
```

```{r complex summary, echo=FALSE, include=FALSE}
summary(model_complex)
```

```{r nested model test with wald test, echo=FALSE, include = FALSE}
# Teacher Attendance
mod2 <- lm(Student.Success~Student.Attendance + Teacher.Attendance, data = my_data)

waldtest(model_naive, mod2, vcov=vcovHC(mod2, type="HC0")) # significant

# Involved Families
mod3 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families , data = my_data)

waldtest(mod2, mod3, vcov=vcovHC(mod3, type="HC0")) # significant

# Supportive Environment
mod4 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment, data = my_data)

waldtest(mod3, mod4, vcov=vcovHC(mod4, type="HC0")) # significant

# Safety
mod5 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety, data = my_data)

waldtest(mod4, mod5, vcov=vcovHC(mod5, type="HC0")) # significant

# Effective Leaders
mod6 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Effective.Leaders, data = my_data)

waldtest(mod5, mod6, vcov=vcovHC(mod6, type="HC0")) # not significant

# Ambitious Instruction
mod7 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Ambitious.Instruction, data = my_data)

waldtest(mod5, mod7, vcov=vcovHC(mod7, type="HC0")) # not significant

# Collaborative Teachers
mod8 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Collaborative.Teachers, data = my_data)

waldtest(mod5, mod8, vcov=vcovHC(mod8, type="HC0")) # not significant

# Discipline Variables
mod9 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Suspension.Days + Suspensions.Percent, data = my_data)

waldtest(mod5, mod9,  vcov=vcovHC(mod9, type="HC0")) # not significant

# years on probation
mod10 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Years.on.Probation, data=my_data)

waldtest(mod5, mod10,  vcov=vcovHC(mod10, type="HC0")) # significant

# latitude
mod11 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Years.on.Probation + Latitude, data=my_data)

waldtest(mod10, mod11,  vcov=vcovHC(mod11, type="HC0")) # not significant

# longitude
mod12 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Years.on.Probation + Longitude, data=my_data)

waldtest(mod10, mod12,  vcov=vcovHC(mod12, type="HC0")) # not significant

# School.Track
mod13 <- lm(Student.Success~Student.Attendance + Teacher.Attendance + Involved.Families + Supportive.Environment + Safety + Years.on.Probation + School.Track, data=my_data)

waldtest(mod10, mod13,  vcov=vcovHC(mod13, type="HC0")) # not significant

```

# Results

Table 1 below compares four models: our "Naive" model with just one predictor variable, $XX$, a "Complex" model which includes all our predictor variables, a "Controllable" model with features that could be reformed by Chicago elementary school superintendents, and a "Controllable Geo" model which includes all features of our "Controllable" model plus the geo location variables, like $Latitude$ and $Longitude$.

```{r Stargazer Table, message=FALSE, echo=FALSE, results='asis'}

stargazer(model_naive, learning_environment, controls, model_complex,
          title = "Model Comparison",
          column.labels = c("Naive", "Learning Environment", "Controls", "Complex"),
          dep.var.caption = "Regression Table",
          dep.var.labels.include = FALSE,
          omit = c("Effective.Leaders", "Ambitious.Instruction", "Collaborative.Teachers", "Involved.Families", "Supportive.Environment", "Safety"),
          add.lines = list(
            c("Involved Families", "", "\\checkmark", "\\checkmark","\\checkmark"),
            c("Supportive Environment", "", "\\checkmark", "\\checkmark", "\\checkmark"),
            c("Safety", "", "\\checkmark","\\checkmark","\\checkmark"),
            c("Effective.Leaders", "","","","\\checkmark"),
            c("Ambitious.Instruction", "","","","\\checkmark"),
            c("Collaborative.Teachers", "","","","\\checkmark"),
            "\\hline"
          ), 
          single.row = TRUE,  # Combine into a single row
          type = "latex", # Output in LaTeX format
          digits=2,
          
          ) 
```

In our naive model, $Growth.Reading.and.Math$ is a significant predictor variable that explains XX% of variability in our model. The percent of variability explain by the model, as indicated by the adjusted r-squared, increases to 74-76% after adding more random variables that help explain the remaining unsystematic variability in the model.

Across all four models, the coefficient of $Growth.Reading.and.Math$ was shown to be significant at a 99% confidence interval. The coefficients of $Student.Attendance$, $Years.on.Probation$, $Involved.Families$, $Safety$, and $Longitude$ were also highly significant (p \< 0.01) in all the models these variables were present in.

# Limitations

Given our large sample size (218 observations), we can utilize a multiple regression model to assess the key drivers on student success, assuming that our data meets the following statistical limitations: The data must be Independent and Identically Distributed (IID), there must exist no perfect collinearity between features.

Each row in our dataframe represents a unique elementary school in Chicago. Because the data includes the entire population of elementary schools in Chicago rather than a specific sample, we have reduced concerns related to sampling bias since every school is represented in the dataset. Since we are analyzing the entire population of elementary schools in Chicago, we assume that the schools operate under similar conditions and follow similar educational standards and policies. This supports the notion that our observations are drawn from the same distribution, therefore meeting the IID assumption. That said, our scope is limited towards only being able to make generalizations about Chicago public schools.

We examined the relationships between independent variables through a correlation matrix and found no correlations near 1, indicating no perfect collinearity. The strongest correlation observed was between $Collaborative.Teachers$ and $Effective.Leaders$ (r = 0.78), which is well below the threshold for concern. We also conducted a Variance Inflation Factor (VIF) test on our complex model, with all VIF values below 5, further confirming the absence of significant multicollinearity.

Since our data is IID and has no perfect collinearity, we pass all of our large linear model assumptions and therefore rule out any concerns for using a multiple regression model for this analysis.

Regarding structural limitations, one key omitted variable is school funding, which could have significant implications for our model's accuracy. School funding is expected to be positively correlated with our target variable, student success. Moreover, more school funding would likely enhance some of our learning environment variables like safety and student attendance. By having school funding as an omitted variable, we may be inadvertently attributing more significance to our existing predictor variables, thus overestimating their impact. Including data on school funding would therefore drive the main effect towards zero, reducing the likelihood of a type I error.

# Conclusion

Through our analysis, we can conclude a student's success in reading and math is significantly impacted by their school attendance, growth, school status, and of course, environmental factors including family involvement, location, and safety.

Since student success can benefit society in many ways, from increasing tax revenue to creating a more vibrant society[^5], it is crucial for superintendents to use this information to inform future roadmaps for success. As a starting point, superintendents can focus on increasing student attendance, which had the largest coefficient amongst our significant predictor variables. As shown in our best-performing model, the "Complex" model, for every percent change in student attendance, student performance on reading and math increases 1.12%. Given these results, Chicago elementary school superintendents could prioritize reforms that help increase student attendance, particularly in southern and eastern parts of the city.

[^5]: How do college graduates benefit society at large?. APLU. (2023, March 1). <https://www.aplu.org/our-work/4-policy-and-advocacy/publicuvalues/societal-benefits/>

That said, addressing the challenge of student attendance requires a multifaceted approach; one that goes beyond simply boosting attendance numbers. It calls for initiatives that foster a safer school environment, cultivate family engagement, and address underlying issues such as safety concerns that may be hindering attendance in the first place. These efforts can resonate beyond attendance, potentially reducing the number of years a school is on probation---another significant predictor of student success in our study.

In conclusion, our study serves as a testament to the power of data-driven decision-making in education. By delving into the complex interplay of factors that contribute to student success, we have provided actionable insights that can be harnessed to foster a more enriching educational experience for elementary school students in Chicago.
